\RequirePackage{snapshot}
\documentclass[final]{ulthesis}

\include{journals}

\usepackage[pdfauthor={Patrick Trainor},
            pdftitle={Bayesian analytical approaches for metabolomics},
            pdfsubject={Bayesian Metabolomics},
            pdfkeywords={Bayesian Statistics, Metabolomics, Mass Spectrometry, Interactome},
            pdfproducer={Latex with hyperref},
            pdfcreator={latex->dvips->ps2pdf},
            pdfpagemode=UseOutlines,
            bookmarksopen=true,
            bookmarksnumbered=true]{hyperref}
\usepackage{memhfixc}

%Additional packages that are useful
\usepackage[pdftex]{graphicx}
\usepackage{verbatim,color,amsmath,amssymb,amsthm,amsfonts,listings,subfig} 
\usepackage{hyperref}  % use without arguments for no metatdata
\usepackage[labelsep=period, labelfont=bf]{caption} % boldface with period
\usepackage{chngcntr}   % use to serially number figures and tables 
\usepackage{algorithm,algorithmic}
\usepackage[paper=portrait,pagesize]{typearea}
\usepackage{setspace}
\usepackage{breakcites}
\usepackage{pdflscape}
\DeclareMathOperator{\PP}{P}
\DeclareMathOperator{\EE}{E}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\ent}{H}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\DE}{DE}
\DeclareMathOperator{\EXP}{EXP}
\DeclareMathOperator{\GA}{GA}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]

\makeatletter
\renewcommand\@memmain@floats{%

% Enable serially numbered figures, tables and equations
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}
\counterwithout{equation}{chapter}}

% Suppress section numbering in TOC
\renewcommand{\cftsectionpresnum}{\begin{lrbox}{\@tempboxa}}
\renewcommand{\cftsectionaftersnum}{\end{lrbox}}
\makeatletter
\setlength{\cftsectionnumwidth}{0pt}

% Restarts section numbered from 1 on each chapter
\renewcommand{\thesection}{\arabic{section}}
\renewcommand\bibname{REFERENCES}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\RequirePackage[left=1.5in,right=1in,top=1in,bottom=1in,footskip=.5in]{geometry} 

\begin{document}
% ---------------------------------------
% Front matter
% ---------------------------------------
%Dissertation author 
\author{Patrick J. Trainor}

% Degree
\authordegree{M.S. in Biostatistics, 2014\\M.A. in Mathematics, 2014\\B.S. in Mathematics, 2012}

% Dissertation title
\title{BAYESIAN ANALYTICAL APPROACHES FOR METABOLOMICS: A NOVEL METHOD FOR MOLECULAR STRUCTURE-INFORMED METABOLITE INTERACTION MODELING, A NOVEL DIAGNOSTIC MODEL FOR DIFFERENTIATING MYOCARDIAL INFARCTION TYPE, AND APPROACHES FOR COMPOUND IDENTIFICATION GIVEN MASS SPECTROMETRY DATA}
%\title{Bayesian analytical approaches for metabolomics: a novel method for molecular structure-informed metabolite interaction modeling, a novel diagnostic model for differentiating Myocardial Infarction type, and  approaches for compound identification given mass spectrometry data}

% Dissertation degree
\dissertationdegree{Doctor of Philosophy}

% Dissertation discipline (may differ from the department name)
\dissertationdiscipline{Interdisciplinary Studies:  Bioinformatics}

% Month and year of degree award
\dissertationdate{August 2018}

% Month day and year of dissertation approval
\approvaldate{July 19, 2018}

% College or school
\school{School of Interdisciplinary and Graduate Studies }

% Department in which the candidate is enrolled
\department{School of Interdisciplinary and Graduate Studies}

% Advisors and committee members
\advisor{Shesh N. Rai, Ph.D.}
\firstmember{Andrew P. DeFilippis, M.D., M.Sc.}
\secondmember{Eric C. Rouchka, D.Sc.}
\thirdmember{Juw W. Park, Ph.D.}
\fourthmember{Fourth Member}

% Use the same keywords for the pdf headers above
\keywords{Bayesian Statistics, Metabolomics, Mass Spectrometry, Interactome}

%This builds the title page and other front content
\frontmatter

\maketitle
% Now we add edited content page by page.
% Retain each section, edit as needed.
% Keep this order of sections and end with the tables of contents.

\begin{dedication}
This dissertation is dedicated to my family: Colleen, Jack, Katie, Winn, Aaron, Matt, Abby, Cassie, Sam, Betty, Jim, Jane, and Edmond; to the ESL students at the BLC; and to the 3\textsuperscript{rd}-5\textsuperscript{th} grade students at the BLC.
\end{dedication}

\begin{acknowledgments}
Acknowledgments are due to the past and present members of the Atherosclerosis / Atherothrombosis Research Laboratory (AARL), especially: Dr. Andrew DeFilippis, Dr. Alok Amraotkar, Dr. Mahrokh Nokhbehzaeim, Amanda Coulter, Allison Smith, Sharon Vincent, Ayesha Singh, Mallory Hatfield, and Dr. Yong Siow. In addition to the laboratory, special thanks are due to the human participants who have graciously agreed to participate in the clinical studies conducted by the AARL to further research into the leading cause of global mortality: heart disease.

Acknowledgments are due to collaborators who have assisted in this work including: Samantha Carlisle, Dr. Hunter Moseley, and Joshua Mitchell.

Acknowledgments are due to Dr. Shesh Rai, Dr. Aruni Bhatnagar, and Dr. Andrew DeFilippis for providing mentorship and support. 

Finally, acknowledgments are due to the following organizations for providing financial support: the American Heart Association (11CRP7300003), the National Institutes of Health
(1P20 GM103492-06 \& Common Fund metabolomics pilot and feasibility award), and the Alpha Phi Foundation (2016 Heart-to-Heart Grant).
\end{acknowledgments}

% Dissertation abstract 
\begin{dissertationabstract}
\begin{center}
	BAYESIAN ANALYTICAL APPROACHES FOR METABOLOMICS: A NOVEL METHOD FOR MOLECULAR STRUCTURE-INFORMED METABOLITE INTERACTION MODELING, A NOVEL DIAGNOSTIC MODEL FOR DIFFERENTIATING MYOCARDIAL INFARCTION TYPE, AND APPROACHES FOR COMPOUND IDENTIFICATION GIVEN MASS SPECTROMETRY DATA \\
	Patrick J. Trainor \\
	July 19, 2018
\end{center}
Metabolomics, the study of small molecules in biological systems, has enjoyed great success in enabling researchers to examine disease-associated metabolic dysregulation and has been utilized for the discovery biomarkers of disease and phenotypic states. In spite of recent technological advances in the analytical platforms utilized in metabolomics and the proliferation of tools for the analysis of metabolomics data, significant challenges in metabolomics data analyses remain. In this dissertation, we present three of these challenges and Bayesian methodological solutions for each. In the first part we develop a new methodology to serve a basis for making higher order inferences in metabolomics, which we define as the testing of hypotheses that are more complex than single metabolite hypothesis tests. This methodology utilizes informative priors that are generated via the analysis of molecular structure similarity to enable the estimation of metabolite ``interactomes'' (or probabilistic models) which are organism-, sample media-, and condition-specific as well as comprehensive; and that can serve as reference models for studying perturbations in metabolic systems. After discussing the development of our methodology, we present an evaluation of its performance conducted using simulation studies, and we use the methodology for estimating a plasma metabolite interactome for stable heart disease. This interactome may serve as a reference model for evaluating systems-level changes that occur with acute disease events such as myocardial infarction (MI) or unstable angina. In the second part of this work, we present the challenge of developing diagnostic classification models which utilize metabolite abundances and that do not ``overfit'' relatively small sample sizes, especially given the high dimensionality of metabolite data acquired using platforms such as liquid chromatography-mass spectrometry. We use a Bayesian methodology for estimating a multinomial logistic regression classifier for the detection and discrimination of the subtype of acute myocardial infarction utilizing metabolite abundance data quantified from blood plasma. As heart disease is the leading cause of global mortality, a blood-based and non-invasive diagnostic test that could differentiate between MI types at the time of the event would have great utility. In the final part of this dissertation we review Bayesian approaches for compound identification in metabolomics experiments that utilize liquid chromatography-mass spectrometry which remains a challenging problem.
\end{dissertationabstract}

% These commands will generate the required tables of contents. 
% The \clearpage commands insure that the next one begins on a new page.
{\newgeometry{top=1in,left=1.5in,bottom=1in,footskip=.5in} \setlength{\beforechapskip}{.75in} 
	\hypersetup{
		colorlinks=false,
		pdfborder={0 0 0},
	}
	\tableofcontents* \clearpage}
{\newgeometry{top=1in,left=1.5in,bottom=1in,footskip=.5in} \setlength{\beforechapskip}{.75in}
	\hypersetup{
		colorlinks=false,
		pdfborder={0 0 0},
	}
	 \listoftables \clearpage}
{\newgeometry{top=1in,left=1.5in,bottom=1in,footskip=.5in} \setlength{\beforechapskip}{.75in}
	\hypersetup{
		colorlinks=false,
		pdfborder={0 0 0},
	}
	 \listoffigures \clearpage}

% ---------------------------------------------
% Main matter
% ---------------------------------------------
% Note we need this to insure proper formatting for the body of the dissertation.
\mainmatter

% Use separate files for chapters and include the chapter titles here.
% Having the titles here is necessary to get them properly indexed and 
% formatted.
\chapter{INTRODUCTION}
\input{intro.tex}

\chapter{PRELIMINARIES: METABOLOMICS}
\input{prelimMetabolomics.tex}

\chapter{PRELIMINARIES: BAYESIAN STATISTICS}
\input{prelimBayesianStats.tex}

\chapter{PRELIMINARIES: GAUSSIAN GRAPHICAL MODELING}
\input{prelimGGM.tex}

\chapter{HIGHER-ORDER INFERENCE IN METABOLOMICS}
\input{HigherOrderIn.tex}

{\newgeometry{top=1in,left=1.5in,bottom=1in,footskip=.5in} 
\chapter{INFERRING METABOLITE INTERACTOMES VIA BAYESIAN GRAPHICAL MODEL SELECTION UTILIZING MOLECULAR STRUCTURE INFORMATIVE PRIORS}
\input{structureAdBLG.tex}
}

{\newgeometry{top=1in,left=1.5in,bottom=1in,footskip=.5in} 
\chapter{A PLASMA INTERACTOME FOR STABLE HEART DISEASE}
\input{heartDiseaseInteractome.tex}
}

\chapter{A BAYESIAN DIAGNOSTIC MODEL FOR DIFFERENTIATING MI TYPE}
\input{diagnostic.tex}

\chapter{BAYESIAN APPROACHES FOR COMPOUND IDENTIFICATION GIVEN LC-MS DATA}
\input{priorCompID.tex}

\chapter{CONCLUSION, DISCUSSION, AND FUTURE DIRECTIONS}
\input{conclusions.tex}

\backmatter
{
\newgeometry{top=1in,bottom=1in,footskip=.5in}
\begin{DoubleSpace*}

\bibliographystyle{apalike}
%\bibliographystyle{plain}
%\bibliographystyle{osajnl}
\bibliography{dissertation}

\end{DoubleSpace*}
% Appendices and the Vita are treated as chapters in the back matter

\newgeometry{top=1in,bottom=1in,footskip=.5in,left=1.5in}{
\chapter{APPENDIX A: ACRONYMS UTILIZED}
\input{acronyms.tex}}

\newgeometry{top=1in,bottom=1in,footskip=.5in,left=1.5in}{
\chapter{APPENDIX B: SOFTWARE DEVELOPED FOR THE BAYESIAN GRAPHICAL LASSO}
\input{code.tex}}

\newgeometry{top=1in,bottom=1in,footskip=.5in,left=1.5in}{
\chapter{CURRICULUM VITAE}
\input{vita.tex}}

\end{document}
