\begin{center}
Patrick Trainor 
\end{center}

\bigskip
{\parindent0pt

\section*{Objective}
My objective is to obtain a research position in computational metabolomics. My primary research interests are: (1) developing Bayesian methodology for integrating extra-experimental knowledge and data sources for integrated -omics analyses, (2) developing Bayesian methodologies for solving problems in analytical chemistry, and (3) developing computational tools for metabolomics data analyses.  

\section*{Education}
 PhD Bioinformatics  \\
University of Louisville, Louisville, KY, 2015-2018 (expected)  \\ \\
 MS Biostatistics  \\
University of Louisville, Louisville, KY, 2012-2014  \\ \\
MA Mathematics  \\
University of Louisville, Louisville, KY, 2012-2014 \\ \\
BS Mathematics  \\
Seattle University, Seattle, WA, 2012

\section*{Research Positions}
\emph{Graduate Research Assistant} \hfill 2012-2014, 2015-Current \\
Institute of Molecular Cardiology \& J.G. Brown Cancer Center (formerly), University of Louisville \\

\emph{Consultant \& Intern} \hfill 2012-2013 \\
Theravance Inc. (Theravance Biopharma) 

\section*{Teaching Positions}
\emph{Graduate Teaching Assistant} \hfill 2012-2014 \\
University of Louisville \\

\emph{Undergraduate Research Assistant} \hfill 2010-2012 \\
Seattle University

\section*{Other Professional Experience}
\emph{Clinical Analytics Consultant} \hfill 2014-2015 \\
Humana, Inc., Louisville, KY \\

\emph{Rifleman \& Infantry Squad Leader} \hfill 2007-2013 \\
4th Anti-Terrorism Battalion \\
United States Marine Corps (R) 

\section*{Publications}
*Denotes equally contributing authorship 

\textbf{Trainor, P.J.}, Yampolskiy, R.V.*, \& DeFilippis, A.P.* (2018). Wisdom of artificial crowds feature selection in untargeted metabolomics: An application to the development of a blood-based diagnostic test for thrombotic myocardial infarction. \emph{Journal of Biomedical Informatics, 81}.  doi: 10.1016/j.jbi.2018.03.007  \\

DeFilippis, A.P., \textbf{Trainor, P.J.} (2018). When given a lemon, make lemonade: Revising cardiovascular risk prediction scores. \emph{Annals of Internal Medicine,} online ahead of print. doi: 10.7326/M18-1175. \\

Carlisle, S.M., \textbf{Trainor, P.J.}, Doll, M.A., Stepp, M.W., Klinge, C.M., \& Hein, D.W. (2018). Knockout of human arylamine N-acetyltransferase 1 (NAT1) in MDA-MB-231 breast
cancer cells leads to increased reserve capacity, maximum mitochondrial capacity, and glycolytic reserve capacity. \emph{Molecular Carcinogenesis,} online ahead of print. doi: 10.1002/mc.22869\\

\textbf{Trainor, P.J.}, DeFilippis, A.P., \& Rai, S.N. (2017). Classifier performance for multiclass phenotype discrimination in metabolomics. \emph{Metabolites, 7}(2). doi: 10.3390/metabo7020030 \\

\textbf{Trainor, P.J.}, Hill, B.G., Carlisle, S.M., Rouchka, E.C., Bhatnagar, A., Rai, S.N., \& DeFilippis, A.P. (2017). Systems characterization of differential plasma metabolome perturbations following thrombotic and non-thrombotic myocardial infarction \emph{Journal of Proteomics, 160}. doi: 10.1016/j.jprot.2017.03.014. \\

DeFilippis, A.P., \textbf{Trainor, P.J.}, Hill, B.G., Amraotkar, A.R., Rai, S.N., Hirsch, G.A., Rouchka, E., \& Bhatnagar, A. (2017). Identification of a plasma metabolomic signature of thrombotic myocardial infarction that is distinct from non-thrombotic myocardial infarction and stable coronary artery disease. \emph{PLoS One, 12}(4). doi: 10.1371/journal.pone.0175591. \\

Gibb, A.A., Epstein, P.N., Uchida, S., McNally, L.A., Obal, D., Katragadda, K., \textbf{Trainor, P.J.},  Conklin, D.J., Brittian, K.R., Tseng, M.T., Wang J., Jones, S.P., Bhatnagar, A., \& Hill, B.G. (2017). Exercise-Induced Changes in Glucose Metabolism Promote Physiologic Cardiac Growth \emph{Circulation, 136}(21). doi: 10.1161/CIRCULATIONAHA.117.028274 \\

Khanal, S. \textbf{Trainor, P.J.}, Zahin, M., Ghim, S.J., Joh, J., Rai, S.N., Jenson, A.B., \& Shumway, B.S. (2017). Histologic variation in high grade oral epithelial dysplasia when associated with high-risk human papillomavirus.  \emph{Oral Surgery, Oral Medicine, Oral Pathology, and Oral Radiology, 123}(5). doi: 10.1016/j.oooo.2017.01.008. \\

Sultan, A., Zheng, Y., \textbf{Trainor, P.J.}, Yong, S., Amraotkar, A.R.  Hill, B.G., \& DeFilippis, A.P. (2017). Circulating prolidase activity in patients with myocardial infarction. \emph{Frontiers in Cardiovascular Medicine}. doi: 10.3389/fcvm.2017.00050.  \\

Amraotkar, A.R., Ghafghazi, S., \textbf{Trainor, P.J.}, Hargis C.W., Irfan A.B., Rai S.N., Bhatnagar A., \& DeFilippis A.P. (2017). Presence of multiple coronary angiographic characteristics for the diagnosis of acute coronary thrombus. \emph{Cardiology Journal, 24}(1). doi: 10.5603/CJ.a2017.0004.  \\

Khosravi, F., \textbf{Trainor, P.J.}, Lambert, C., Kloecker, G., Wickstrom, E. Rai, S.N., \& Panchapakesan, B. (2016). Static micro-array isolation, dynamic time series classification, capture and enumeration of spiked breast cancer cells in blood: the nanotube--CTC chip. \emph{Nanotechnology, 27}(44). doi: 10.1088/0957-4484/27/44/44LT03.  \\

Carlisle, S.M., \textbf{Trainor, P.J.}, Yin, X., Doll, M.A., Stepp, M.W., States, J.C., Zhang, X., \& Hein, D.W. (2016). Untargeted polar metabolomics of transformed MDA-MB-231 breast cancer cells expressing varying levels of human arylamine N-acetyltransferase 1. \emph{Metabolomics, 12}(7). doi: 10.1007/s11306-016-1056-z.  \\

Amraotkar, A.R., Song, D.D., \textbf{Trainor, P.J.}, Ismail, I., Kothari, K., Sing, A., Moore, J.B., Rai, S.N., Bhatnagar, A., \& DeFilippis, A.P. (2016). Platelet count and mean platelet volume at the time of and after acute myocardial infarction. \emph{Clinical and Applied Thrombosis/Haemostasis, 23}(8). doi: 10.1177/1076029616683804.  \\

DeFilippis, A.P., Chernyavskiy, I., Amraotkar, A.R., \textbf{Trainor, P.J.}, 
Kothari, S., Ismail, I., Hargis, C.W., Korley, F.K., Tsimikas, S., Rai, S.N., \& Bhatnagar, A. (2016). Circulating levels of plasminogen and oxidized phospholipids bound to plasminogen distinguish between atherothrombotic and non-atherothrombotic myocardial infarction. \emph{Journal of Thrombosis and Thrombolysis, 42}(1). doi: 10.1007/s11239-015-1292-5.  \\

Egan, J.M., Sloughter, J.M., Cardoso, T., \textbf{Trainor, P.J.}, Wu, K., Safford, H., \& Fournier, D. (2016). Multi-temporal ecological analysis of Jeffrey pine beetle outbreak dynamics within the Lake Tahoe Basin. \emph{Population Ecology, 58}(3). doi: 10.1007/s10144-016-0545-2.  \\

Rai, S.N.*, \textbf{Trainor, P.J.}*, Khosravi, F., \& Panchapakesan, B.  (2016). Classification of biosensor time series using dynamic time warping: applications in screening cancer cells with characteristic biomarkers. \emph{Open Access Medical Statistics, 6}. doi: 10.2147/OAMS.S104731.  \\

Khosravi, F., \textbf{Trainor, P.J.}, Rai, S.N., Kloecker, G., Wickstrom, E. \& Panchapakesan, B. (2016). Label-free capture of breast cancer cells spiked in buffy coats using carbon nanotube antibody micro-arrays. \emph{Nanotechnology, 27}(13). doi: 10.1088/0957-4484/27/13/13LT02.  \\

Khanal, S., Cole, ET., Joh, J., Ghim, S.J., Jenson, A.B., Rai, S.N., \textbf{Trainor, P.J.}, \& Shumway, B.S. (2015). Human papillomavirus detection in histologic samples of multifocal epithelial hyperplasia: a novel demographic presentation. \emph{Oral Surgery, Oral Medicine, Oral Pathology, and Oral Radiology, 120}(6). doi: 10.1016/j.oooo.2015.07.035. 

\section*{Published Software}
\textbf{Trainor, P.J.} \& Wang H. (2017). \emph{BayesianGLasso: Bayesian Graphical Lasso}. Available from https://CRAN.R-project.org/package=BayesianGLasso

\section*{Accepted Publications}
Khanal S., Shumway B.S., Zahin M., Redman R., \textbf{Trainor P.J.}, Rai S.N., Ghim S.J., Jenson A.B., \& Joh J. (2018). Human Papillomavirus DNA integration, but not viral methylation, is associated with malignant progression of head and neck cancers. \emph{Oncotarget}, in production.

\section*{Invited Talks}
\textbf{Trainor, P.J.} (2017). Metabolomics of thrombotic myocardial infarction: systems characterization of plasma metabolome perturbations and the development of a diagnostic classifier. \emph{Systems Biology Seminar Series}, University of Kentucky, Lexington, Kentucky.

\section*{Oral / Poster Presentations}
\textbf{Trainor, P.J.} (2018). Inferring metabolite interactomes via Bayesian graphical model selection utilizing molecular structure informative priors. \emph{14\textsuperscript{th} Annual Conference of the Metabolomics Society}, Seattle, Washington.  \\ 

\textbf{Trainor, P.J.} (2018). High Performance R: Parallel computing, seamless C++ integration, optimizing linear algebra routines, and deep learning in the R language. \emph{Kentucky Biomedical Research Infrastructure Network Seminar Series}, University of Louisville, Louisville, Kentucky.  \\ 

\textbf{Trainor, P.J.}, Carlisle, S.M., Hill, B.G., Rouchka, E.C., Bhatnagar, A., Rai, S.N., \& DeFilippis, A.P. (2017). Metabolomic analysis of thrombotic and non-thrombotic Myocardial Infarction: From systems biology to diagnostic classification. \emph{Symposium in Molecular Biology ``Metabolism: Disease Models and Model Organisms''}. Pennsylvania State University, State College, Pennsylvania.  \\ 

\textbf{Trainor, P.J.}, Carlisle, S.M., DeFilippis, A.P.*, \& Rai, S.N.* (2017). Molecular fingerprinting for inferring a Gaussian Graphical Model representation of a stable coronary artery disease plasma interactome using adaptive graphical Lasso penalization. \emph{UT-KBRIN Annual Bioinformatics Summit}, Burns, Tennessee.  \\ 

\textbf{Trainor, P.J.}, Bhatnagar, A., Rai, S.N., \& DeFilippis, A.P. (2016). Dynamic changes in topologically related plasma metabolites following atherothrombotic and non-atherothrombotic myocardial infarction: A weighted network analysis. \emph{Kentucky Academy of Science Annual Meeting}, University of Louisville, Louisville, Kentucky.  \\ 

\textbf{Trainor, P.J.}, \& Rai, S.N. (2015). Comparison of partial least squares-discriminant analysis, random forests, and rotation forests for phenotype discrimination given metabolomic data. \emph{Research!Louisville}, University of Louisville, Louisville, Kentucky.  \\ 

\textbf{Trainor, P.J.} \& Rai, S.N. (2013). Patient rule induction method for identifying subgroups in clinical studies. \emph{Kentucky Academy of Science Annual Meeting}, Morehead State University, Morehead, Kentucky. \\ 

\textbf{Trainor, P.J.}, \& Barnes, C. (2013). Statistical considerations for dedicated cardiovascular safety studies. \emph{Theravance, Inc. (now Theravance Biopharma) company seminar}, San Francisco, California. \\ 

\textbf{Trainor, P.J.}, \& Sloughter, J.M. (2012). Generalized Linear Modeling in Population Ecology. \emph{Northwest Undergraduate Mathematics Symposium}, Lewis \& Clark College, Portland, Oregon. 

\section*{Grants}
 Co-investigator. NIH West Coast Metabolomics Center Pilot and Feasibility Award: ``Targeted plasma metabolomic profiling of thrombotic myocardial infarction''. PI: DeFilippis, A.P. (2017-2018). \$50,000 direct cost. \\

 Co-Investigator. Alpha Phi Foundation Heart-to-Heart Grant (2016 award): ``Circulating Levels of Oxidized Phospholipids, Sub-Clinical Atherosclerosis and Atherosclerotic Cardiovascular Disease''. PI: DeFilippis, A.P. (2016-2017). \$100,000 direct costs. 

\section*{Computing Experience} 
\emph{Programming Languages:} R, Python, C++, MATLAB/Octave \\
\emph{Machine Learning \& Deep Learning:} TensorFlow, Keras, WEKA \\
\emph{Statistical Software:} R, SAS \\
\emph{Bioinformatics Software:} Many packages distributed via Bioconductor \\
\emph{Computer Algebra Systems:} SAGE, Mathematica \\ 
\emph{Operating Systems:} UNIX, Linux (openSUSE is my favorite flavor), and Windows

\section*{AWARDS}
Metabolomics Society Student Travel Award, 2018, Metabolomics Society\\

2\textsuperscript{nd} Place Graduate Research Competition in Health Sciences, 2016, Kentucky Academy of Science\\

1\textsuperscript{st} Place Graduate Research Competition in Mathematics, 2013, Kentucky Academy of Science\\

Janet E. Mills Award for Outstanding Undergraduate Research in Mathematics, Seattle University

}